---
title: "Module 5 Project: Using Linear Programming Models to maximize profits"
author: "Thuy Nhu Thao Tran"
output: 
  pdf_document: 
    latex_engine: xelatex
---

# Introduction

# 1. Mathematical formulation of the problem

First of all, I calculated **the total warehouse space** is **2460ft** because:

\- The warehouse has 82 shelves.

\- Each shelf is 30 ft long and 5 ft wide.

-\> 82 x 30 = 2460ft

I define decision variables:

\- x1: Number of Pressure Washers purchased

\- x2: Number of Go Karts purchased

\- x3: Number of Generators purchased

\- x4: Number of Water Pump cases purchased

Objective Function (Maximize Net Profit):

maxZ = (499.99x1 + 729.99x2 + 700.99x3 + 269.99x4)−(330x1 + 370x2 + 410x3 + 635x4)

**-\> maxZ = 169.99x1 + 359.99x2 + 290.99x3 − 365.01x4**

Constraints:

-   Budget constraint (Cannot exceed \$170,000): 330x1 + 370x2 + 410x3 + 635x4 =\<170000

-   Warehouse space constraint (Total available: 2460 ft): 5x1 + 8x2 + 5x3 + 1.25x4 =\< 2460

-   Inventory allocation constraint (At least 30% must be pressure washers and go-karts):

    x1 + x2 \>= 0.3(x1 + x2 + x3 + x4)

-   Sales ratio constraint (Generators must be at least twice the water pumps): x3 \>= 2x4

-   Non-negativity constraints: x1, x2, x3, x4 \>= 0

# 2. Set up the linear programming formulation in R, 3. Use R to solve the problem, and generate a sensitivity report.

Then, I use lpSolve package in R to set up and solve the model.

```{r}
# Load the required package
library(lpSolve)

# Define cost and revenue
costs <- c(330, 370, 410, 635)
revenues <- c(499.99, 729.99, 700.99, 269.99)

# Profit per unit
profit <- revenues - costs

# Constraints
## Budget constraint
budget <- 170000

## Warehouse space constraint (Total available: 2460 ft)
space_per_item <- c(5, 8, 5, 5/4)  # Water pumps case occupies 1.25 ft
warehouse_capacity <- 2460

## Inventory allocation constraint (30% allocated to pressure washers and go-karts)
## x1 + x2 >= 0.3 * (x1 + x2 + x3 + x4)

## Sales ratio constraint (Generators must be at least 2 times water pumps)
## x3 >= 2 * x4

# Coefficients matrix
constraints <- matrix(
  c(330, 370, 410, 635,      # Budget constraint
    5, 8, 5, 5/4,           # Warehouse space constraint
    0.7, 0.7, -0.3, -0.3,   # Inventory allocation
    0, 0, 1, -2),           # Sales ratio constraint
  nrow = 4, byrow = TRUE
)

# Right-hand side values for constraints
rhs <- c(budget, warehouse_capacity, 0, 0)

# Constraint directions
constraints_dir <- c("<=", "<=", ">=", ">=")

# Solve the LP model
solution <- lp(
  direction = "max",  # Maximization problem
  objective.in = profit,
  const.mat = constraints,
  const.dir = constraints_dir,
  const.rhs = rhs,
  all.int = TRUE  # Ensure integer values
)

# Display results
cat("Optimal Solution:")
print(solution$solution)  # Number of items to purchase
cat("\nMaximum Profit: ", solution$objval)

```

# 4. Describe the optimal solutions

With the output above, we can see that:

\- Optimal Solution:

-   Pressure Washers (x1): 1 unit

-   Go-Karts (x2): 125 units

-   Generators (x3): 291 units

-   Water Pump Cases (x4): 0 units

\- Maximum Profit: 129846.8

**Why is x4 zero?** Answer: A zero value in a linear programming solution means it is not profitable or not feasible to purchase water pumps. This could be due to:

1.  Low profitability

-   Water pumps have a negative profit per unit: 269.99 - 635 = -365.01
-   Since selling water pumps causes a loss, the model does not include them.

2.  Budget Constraints:

-   Higher-profit items could make greater use of the budget or warehousing.
-   Due to their larger revenues, the model gives priority to go-karts and generators.

3.  Impact of Sales Ratio Constraint:

-   Selling water pumps compels us to have additional generators on hand since generators must be at least twice as large as water pumps.
-   Since this might waste money and space, it would be best to fully avoid using water pumps.

# 5. Use the sensitivity report to determine the smallest selling price

```{r}
# Sensitivity Analysis (Shadow Prices and Reduced Costs)
print(solution$reducedcost)
```

Because the output I got is NULL, I will calculate the smallest selling price manually. Then, I used the table below to compare x4 with the most profitable item (generators, x3) to estimate the required price.

| Item | Cost | Selling price | Profit  |
|------|------|---------------|---------|
| x3   | 410  | 700.99        | 290.99  |
| x4   | 635  | 269.99        | -365.01 |

**Estimate Minimum Selling Price for x4:**

New Selling Price of x₄ = Cost + Profit of Most Profitable Item

= 635 + 290.99 = 925.99

```{r}
revenues[4] <- 925.99  # Increase water pump price
profit <- revenues - costs  # Recalculate profit per unit
```

Re-run the LP model:

```{r}
solution <- lp(
  direction = "max",
  objective.in = profit,
  const.mat = constraints,
  const.dir = constraints_dir,
  const.rhs = rhs,
  all.int = TRUE
)

print(solution$solution)  # Check if x₄ is now included
cat("\nNew Maximum Profit: ", solution$objval)

```

# 6.

```{r}
print(solution$duals)
```

Since the shadow price is 0, this tells us:

\- The company already has enough budget to buy the most profitable combination of products.

\- More money won’t change the optimal solution because the company cannot store or sell more products under the current constraints.

# 7.

Since the shadow price is 0, this tells us:

\- The company's potential to optimize profit is not limited by warehouse area.

\- Changing the warehouse's size will not change the optimal solution or profit.

\- It is not necessary for the company to lease a bigger warehouse.
